---
title: "Okinawa guppies"
author: "Sasha Mikheyev"
date: "10/30/2018"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    theme: cerulean
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = 'index.html') })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmerTest)
library(tidyverse)
library(car)
```

# Site structure
```{r}
mds <- read_table("data/all.mds") %>%
  separate(FID, c("site", "ind"), extra = "merge") %>%
  select(site, id = IID, C1, C2)
ggplot(mds, aes(C1, C2, color = site, label = id)) +
  geom_point() +
  scale_color_brewer(type = "qual") +
  theme_bw()
mds %>% filter(C1 < -0.05 & C2 < -0.05) %>%
  ggplot(aes(C1,C2,label=id, color=site)) +
  geom_text()
```

Looks like there are `site1_male23` and `site1_male55` individuals may have been switched, and we'll drop them just in case

# Read phenotypic data from VCFtools
filtered using vcftools ` --mac 10 --max-missing 0.5 --minQ 20 --minDP 6 `
```{r heterozygosity}
pheno <- read_tsv("data/guppy pheno.tsv", col_types = cols()) %>%
  filter(!Label %in% c("site1_male23", "site1_male55") & TopBottom != "0") %>%
  mutate(orange = rowMeans(cbind(L_Orangearea, R_Orangearea), na.rm = T), site = factor(Site)) %>%
  select(site, Label, orange, TopBottom) %>%
  left_join(read_csv("data/geno2pheno.csv", col_types = cols()), by = c("Label" = "pheno")) %>%
  left_join(read_tsv("data/out.het", col_types = cols()) %>% 
  mutate(het = 1 - `O(HOM)` / N_SITES) %>% 
  select(gen = INDV, het, F, nsites = N_SITES)) %>%
  na.omit()
ggplot(pheno, aes(het, orange, color = site)) +
  geom_point()
with(pheno, cor.test(F, het))

ggplot(pheno, aes(site, het)) +
  geom_boxplot()

pheno_mod <- glm(het ~ TopBottom * site, data = pheno)
summary(pheno_mod)
Anova(pheno_mod)
```

It appears that the effect of heterozygosity is site-specific. It is strongest in the site with the most variability in heterozygosite (Site 3)

# Fluctuating asymmetry

```{r flucAsym}
phenoFA <- read_tsv("data/guppy pheno.tsv", col_types = cols()) %>%
  filter(!Label %in% c("site1_male23", "site1_male55") & TopBottom != "0") %>%
  mutate(blackFA = L_Blackarea - R_Blackarea, orangeFA = L_Orangearea - R_Orangearea, site = factor(Site)) %>%
  select(site, Label, blackFA, orangeFA, TopBottom) %>%
  left_join(read_csv("data/geno2pheno.csv", col_types = cols()), by = c("Label" = "pheno")) %>%
  left_join(read_tsv("data/out.het", col_types = cols()) %>% mutate(het = 1 - `O(HOM)` / N_SITES) %>% select(gen = INDV, het, F, nsites = N_SITES)) %>%
  na.omit() %>%
  group_by(site) %>%
  transmute(het = het, absOrangeFA = abs(orangeFA), meanSite = mean(het), withinSite = het - meanSite, absBlackFA = abs(blackFA)) %>%
  ungroup()
hist(phenoFA$absOrangeFA)
hist(phenoFA$absBlackFA)
modelFA.sameSlopes <- glmer(round(absOrangeFA) ~ het + meanSite + (1 | site), data = phenoFA, family = poisson) # https://www.statalist.org/forums/forum/general-stata-discussion/general/1443819-dependent-variable-in-absolute-values
summary(modelFA.sameSlopes)

modelFA.sameSlopes <- glmer(round(absBlackFA) ~ het + meanSite + (1 | site), data = phenoFA, family = poisson) # https://www.statalist.org/forums/forum/general-stata-discussion/general/1443819-dependent-variable-in-absolute-values
summary(modelFA.sameSlopes)
```

